Ethereum: A Secure Decentralised Generalised Transaction Ledger

### 2014-04-07 Paper.tex
Formal stuff mostly complete.

\definecolor{lightyellow}{rgb}{1,0.98,0.9}

@@

\pagecolor{lightyellow}

@@

and $f$ is the Ethereum state transition function. In Ethereum, $\Upsilon$, together with $\mathbb{S}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\mathbb{S}$ allows components to store arbitrary state between transactions.

@@

\begin{eqnarray}
\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, \mathbf{T}) \\
B & \equiv & (..., [ T_0, T_1, ... ]) \\
\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
\end{eqnarray}

@@

This is the basis of the blockchain paradigm, a model that forms the backbone of Ethereum.

@@

In order to incentivise computation within the network, there needs to be an agreed method for transmitting value. To address this issue, Ethereum has an intrinsic currency, Ether, known also as ETH. The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted is the Wei. One Ether is defined as being $10^{18}$ Wei. There exist other subdenominations of Ether:

\par\bigskip
\begin{center}
\begin{tabular}{lr}
\toprule
Multiplier & Name \\
\midrule
$10^0$ & Wei \\
$10^{12}$ & Szabo \\
$10^{15}$ & Finney \\
$10^{18}$ & Ether \\
\bottomrule
\end{tabular}
\end{center}
\par

Throughout the document, any reference to value, in the context of Ether, currency, a balance or a payment, should assumed to be counted in Wei.

@@

Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block the resultant structure is necessarily a tree of blocks. In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme. If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.

@@

The world state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP). Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}). The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database. This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state. Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly. Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.

@@

Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\mathbb{S}$ the state:

\begin{equation}
\mathbb{S}' = \Upsilon(\mathbb{S}, T)
\end{equation}

@@

We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:

\begin{eqnarray}
g_0 & \equiv & \begin{cases}
  (|T_\mathbf{i}| + |T_\mathbf{b}|) G_{data} + G_{create} & \text{if} \quad T_r = 0 \\
  |T_\mathbf{d}| G_{data} + G_{call} & \text{otherwise} \\
\end{cases}
\end{eqnarray}

@@

The up-front cost $v_0$ is calculated as:

\begin{equation}
v_0 \equiv T_g T_p + T_v
\end{equation}

the validity is determined as:

\begin{eqnarray}
S(T) & \neq & \varnothing \\ \wedge \quad \mathbb{S}[S(T)] & \neq & \varnothing \\ \wedge \quad T_n & = & \mathbb{S}[S(T)]_{nonce} \\ \wedge \quad g_0 & \leq & T_g \\ \wedge \quad v_0 & \leq & \mathbb{S}[S(T)]_{balance}
\end{eqnarray}

@@

We define the rollback state $\mathbb{S}_0$:

\begin{eqnarray}
\mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - c \\
\mathbb{S}_0[S(T)]_{nonce} & \equiv & \mathbb{S}[S(T)]_{nonce} + 1
\end{eqnarray}

Evaluating $\mathbb{S}_0$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):

\begin{equation}
(\mathbb{S}_P, g') \equiv \begin{cases}
\Lambda(\mathbb{S}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_r = 0 \\
\Theta^*(\mathbb{S}_0, S(T), T_o, T_r, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
\end{cases}
\end{equation}

where

\begin{equation}
g \equiv T_g - g_0
\end{equation}

Note we use $\Theta^*$ to take only the first two components of the function's value; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate and the Ether for the gas actually used is given to the miner, whose address is specified as the coinbase of the present block $B$. So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:

\begin{eqnarray}
\mathbb{S}' & \equiv & \mathbb{S}_o \quad \text{except} \\
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_o[s]_{balance} + g' T_p \\
\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_o[m]_{balance} + (g - g') T_p \\
m & \equiv & {B_H}_b
\end{eqnarray}

@@

The EVM is a simple stack-based architecture. The word size of the machine (and thus size of stack item) is 256-bit. This was chosen to facilitate the SHA3-256 hash scheme and ECC. The memory model is a simple word-addressed byte array. The stack has an unlimited size. The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array. Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state in the state database. All locations in both storage and memory are well-defined initially as zero.

@@

\begin{itemize}
\item {\small SHA3} costs $G_{sha3}$ gas
\item {\small SLOAD} costs $G_{sload}$ gas
\item {\small BALANCE} costs $G_{balance}$ gas
\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
\begin{itemize}
\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
\item $d = 1$ otherwise.
\end{itemize}
\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of contract body code.
\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of contract initialisation code.
\item All other operations cost $G_{step}$ gas.
\end{itemize}

@@

However, instructions do typically alter one or several components of these values. Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$.

@@

Thus we define the total difficulty of block $B$ recursively as:

\begin{eqnarray}
B_t & \equiv & B'_t + B_d + \sum\limits_{U \in B_\mathbf{U}} U_d \\
B' & \equiv & P(H) \\
H & \equiv & B_H \\
\end{eqnarray}

@@

\begin{enumerate}
\item Validate (or, if mining, determine) uncles.
\item Apply rewards.
\item Verify (or, if mining, compute a valid) nonce.
\end{enumerate}

@@

The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of uncle to the present block. Formally:

\begin{equation}
\bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; P(U) = P(P(B_H)) \; \wedge \; P(B_H) \neq B
\end{equation}

@@

The application of rewards to a block involves placing in the account of the miner-determined coinbase address additional Ether according to

@@

For the PoC series, we use a simplified proof-of-work. This is not ASIC resistant and is meant merely as a placeholder. It utilises the SHA3 hash function to

It is formally defined as $\mathcal{P}$:

\begin{equation}
\mathcal{P}(H, n) \equiv \mathcal{B}(\mathcal{S}(\mathcal{S}(\mathcal{R}(H)) \circ n))
\end{equation}

@@

>>>>

\definecolor{lightyellow}{rgb}{1,0.98,0.9}
\definecolor{lightpink}{rgb}{1,0.94,0.95}

@@

\pagecolor{lightyellow}
%\pagecolor{lightpink}

@@

where $\Upsilon$ is the Ethereum state transition function. In Ethereum, $\Upsilon$, together with $\mathbb{S}$ are considerably more powerful then any existing comparable system; $\Upsilon$ allows components to carry out arbitrary computation, while $\mathbb{S}$ allows components to store arbitrary state between transactions.

@@

\begin{eqnarray}
\mathbb{S}_{t+1} & \equiv & \Pi(\mathbb{S}_t, B) \\
B & \equiv & (..., [ T_0, T_1, ... ]) \\
\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...)
\end{eqnarray}

@@

This is the basis of the blockchain paradigm, a model that forms the backbone of not only Ethereum, but all decentralised consensus-based transaction systems to date.

@@

In order to incentivise computation within the network, there needs to be an agreed method for transmitting value. To address this issue, Ethereum has an intrinsic currency, Ether, known also as {\small ETH}. The smallest subdenomination of Ether, and thus the one in which all integer values of the currency are counted, is the Wei. One Ether is defined as being $10^{18}$ Wei. There exist other subdenominations of Ether:

\par\bigskip
\begin{center}
\begin{tabular}{rl}
\toprule
Multiplier & Name \\
\midrule
$10^0$ & Wei \\
$10^{12}$ & Szabo \\
$10^{15}$ & Finney \\
$10^{18}$ & Ether \\
\bottomrule
\end{tabular}
\end{center}
\par

Throughout the present work, any reference to value, in the context of Ether, currency, a balance or a payment, should be assumed to be counted in Wei.

@@

Since the system is decentralised and all parties have an opportunity to create a new block on some older pre-existing block, the resultant structure is necessarily a tree of blocks. In order to form a consensus as to which path, from root (the genesis block) to leaf (the block containing the most recent transactions) through this tree structure, known as the blockchain, there must be an agreed-upon scheme. If there is ever a disagreement between nodes as to which root-to-leaf path down the block tree is the `best' blockchain, then a \textit{fork} occurs.

@@

The world state (\textit{state}, see Appendix \ref{app:state}), is a mapping between addresses (160-bit identifiers) and account states (a data structure serialised as RLP, see Appendix \ref{app:rlp}). Though not stored on the blockchain, it is assumed that the implementation will maintain this mapping in a modified Merkle Patricia tree (\textit{trie}, see Appendix \ref{app:trie}). The trie requires a simple database backend that maintains a mapping of bytearrays to bytearrays; we name this underlying database the state database. This has a number of benefits; firstly the root node of this structure is cryptographically dependent on all internal data and as such its hash can be used as a secure identity for the entire system state. Secondly, being an immutable data structure, it allows any previous state (whose root hash is known) to be recalled by simply altering the root hash accordingly. Since we store all such root hashes in the blockchain, we are able to trivially revert to old states.

@@

Formally, we consider the function $\Upsilon$, with $T$ being a transaction and $\mathbb{S}$ the state:
\begin{equation}
\mathbb{S}' = \Upsilon(\mathbb{S}, T)
\end{equation}

@@

We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{equation}
g_0 \equiv \begin{cases}
  (|T_\mathbf{i}| + |T_\mathbf{b}|) G_{data} + G_{create} & \text{if} \quad T_r = 0 \\
  |T_\mathbf{d}| G_{data} + G_{call} & \text{otherwise} \\
\end{cases}
\end{equation}

@@

The up-front cost $v_0$ is calculated as:
\begin{equation}
v_0 \equiv T_g T_p + T_v
\end{equation}

the validity is determined as:
\begin{eqnarray}
S(T) & \neq & \varnothing \\ \wedge \quad \mathbb{S}[S(T)] & \neq & \varnothing \\ \wedge \quad T_n & = & \mathbb{S}[S(T)]_{nonce} \\ \wedge \quad g_0 & \leq & T_g \\ \wedge \quad v_0 & \leq & \mathbb{S}[S(T)]_{balance}
\end{eqnarray}

@@

We define the rollback state $\mathbb{S}_0$:
\begin{eqnarray}
\mathbb{S}_0 & \equiv & \mathbb{S} \quad \text{except:} \\
\mathbb{S}_0[S(T)]_{balance} & \equiv & \mathbb{S}[S(T)]_{balance} - c \\
\mathbb{S}_0[S(T)]_{nonce} & \equiv & \mathbb{S}[S(T)]_{nonce} + 1
\end{eqnarray}

Evaluating $\mathbb{S}_0$ from $\mathbb{S}_0$ depends on the transaction type; either contract creation or message call; we define the pairing of post-execution provisional state ($\mathbb{S}_P$) and remaining gas ($g'$):
\begin{equation}
(\mathbb{S}_P, g') \equiv \begin{cases}
\Lambda(\mathbb{S}_0, S(T), T_o, g, T_p, T_v, T_\mathbf{i}, T_\mathbf{b}) & \text{if} \quad T_r = 0 \\
\Theta^*(\mathbb{S}_0, S(T), T_o, T_r, g, T_p, T_v, T_\mathbf{d}) & \text{otherwise}
\end{cases}
\end{equation}

where
\begin{equation}
g \equiv T_g - g_0
\end{equation}

Note we use $\Theta^*$ to denote the fact that only the first two components of the function's value are taken; the third represents the message-call's output value (a byte array) and is unused in the context of transaction evaluation.

After the message call or contract creation is processed, the state is finalised by refunding $g'$, the remaining gas, to the sender at the original rate. The Ether for the gas that was actually used is given to the miner, whose address is specified as the coinbase of the present block $B$. So we define the final state $\mathbb{S}'$ in terms of the provisional state $\mathbb{S}_P$:
\begin{eqnarray}
\mathbb{S}' & \equiv & \mathbb{S}_P \quad \text{except} \\
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g' T_p \\
\mathbb{S}'[m]_{balance} & \equiv & \mathbb{S}_P[m]_{balance} + (g - g') T_p \\
m & \equiv & {B_H}_b
\end{eqnarray}

In the case that $s = m$ then we simply return the Ether back to the sender/miner, collapsing the exception into:
\begin{eqnarray}
\mathbb{S}'[s]_{balance} & \equiv & \mathbb{S}_P[s]_{balance} + g
\end{eqnarray}

@@

The EVM is a simple stack-based architecture. The word size of the machine (and thus size of stack item) is 256-bit. This was chosen to facilitate the SHA3-256 hash scheme and elliptic-curve computations. The memory model is a simple word-addressed byte array. The stack has an unlimited size. The machine also has an independent storage model; this is similar in concept to the memory but rather than a byte array, it is a word-addressable word array. Unlike memory, which is volatile, storage is non volatile and is maintained as part of the system state in the state database. All locations in both storage and memory are well-defined initially as zero.

@@

\begin{itemize}
\item {\small SHA3} costs $G_{sha3}$ gas
\item {\small SLOAD} costs $G_{sload}$ gas
\item {\small BALANCE} costs $G_{balance}$ gas
\item {\small SSTORE} costs $d.G_{sstore}$ gas where:
\begin{itemize}
\item $d = 2$ if the new value of the storage is non-zero and the old is zero;
\item $d = 0$ if the new value of the storage is zero and the old is non-zero;
\item $d = 1$ otherwise.
\end{itemize}
\item {\small CALL} costs $G_{call}$, though additional gas may be taken for the execution of contract body code.
\item {\small CREATE} costs $G_{create}$, though additional gas may be taken for the execution of contract initialisation code.
\item {\small STOP} costs $G_{stop}$ gas
\item All other operations cost $G_{step}$ gas.
\end{itemize}

@@

However, instructions do typically alter one or several components of these values. Altered components listed by instruction are noted in Appendix \ref{app:vm}, alongside values for $\alpha$ and $\delta$ and formal a description of the gas requirements.

@@

Thus we define the total difficulty of block $B$ recursively as:
\begin{eqnarray}
B_t & \equiv & B'_t + B_d + \sum\limits_{U \in B_\mathbf{U}} U_d \\
B' & \equiv & P(H) \\
H & \equiv & B_H \\
\end{eqnarray}

@@

\begin{enumerate}
\item Validate (or, if mining, determine) uncles.
\item Apply rewards.
\item Verify (or, if mining, compute a valid) state and nonce.
\end{enumerate}

@@

The validation of uncle headers means nothing more than verifying that each uncle header is both a valid header and satisfies the relation of uncle to the present block. Formally:
\begin{equation}
\bigwedge_{U \in B_\mathbf{U}} V(U) \; \wedge \; P(U) = P(P(B_H)) \; \wedge \; P(B_H) \neq B
\end{equation}

@@

The application of rewards to a block involves raising the balance of the accounts of the coinbase address of the block and each uncle by a certain amount. We raise the block's coinbase account by $R_b$, the block reward, and the coinbase of each uncle by $\frac{7}{8}$ of that. Formally we define the function $\Omega$:
\begin{eqnarray}
\Omega(B, \mathbb{S}) & \equiv & \mathbb{S}': \mathbb{S}' = \mathbb{S} \quad \text{except:} \\
\mathbb{S}'[{B_H}_b] & = & \mathbb{S}[{B_H}_b] + R_b \\
\forall U \in B_\mathbf{U}: \quad \mathbb{S}'[U_b] & = & \mathbb{S}[U_b] + \frac{7}{8}R_b
\end{eqnarray}

We define the block reward as 1500 Finney:
\begin{equation}
\text{Let} \quad R_b = 1.5^{16}
\end{equation}

\subsubsection{State \& Nonce Validation}

We may now define the function, $\Gamma$, that maps a block $B$ to its initiation state:
\begin{equation}
\Gamma(B) \equiv \begin{cases}
\mathbb{S}_0 & \text{if} \quad P(B_H) = \varnothing \\
\mathbb{S}_i: r(\mathbb{S}_i) = {P(B_H)_H}_r & \text{otherwise}
\end{cases}
\end{equation}

Here, $r(\mathbb{S}_i)$ means the hash of the root node of a trie of state $\mathbb{S}_i$; this is stored in the state database since the trie is a mutable data structure.

And finally define $\Phi$, the block transition function, which maps an incomplete block $B$ to a complete block $B'$:
\begin{eqnarray}
\Phi(B) & \equiv & B': \quad B' = B^* \quad \text{except:} \\
B'_n & = & n: \quad \mathcal{P}(B^*, n) < \frac{2^{256}}{H_d} \\
B^* & \equiv & B \quad \text{except:} \quad B'_r = r(\Pi(\Gamma(B), B))
\end{eqnarray}

As specified at the beginning of the present work, $\Pi$ is the state-transition function, which is defined in terms of $\Omega$, the block finalisation function and $\Upsilon$, the transaction-evaluation function, both now well-defined:
\begin{eqnarray}
\Pi(\mathbb{S}, B) & \equiv & \Omega(B, \Upsilon(\Upsilon(\mathbb{S}, T_0), T_1) ...) \\
B & \equiv & (H, [ T_0, T_1, ... ], \mathbf{U})
\end{eqnarray}

Thus the complete block-transition mechanism, less $\mathcal{P}$, the proof-of-work function is defined.

@@

For the PoC series, we use a simplified proof-of-work. This is not ASIC resistant and is meant merely as a placeholder. It utilises the bare SHA3 hash function to secure the block chain by requiring the SHA3 hash of the concatenation of the nonce and the header's SHA3 hash to be sufficiently low.

It is formally defined as $\mathcal{P}$:
\begin{equation}
\mathcal{P}(H, n) \equiv \mathcal{B}(\mathcal{S}(\mathcal{S}(\mathcal{R}(H)) \circ n))
\end{equation}

@@

State database won't keep everything; maintains an age and eventually throws out nodes that are not recent enough or checkpoints.

Throw out nodes in state that haven't sent/received a transaction in over X blocks.
